### 2016-05-22 paper/intro.tex
Minor typos (more to come)

Parameter estimation/inference is an important component of mathematical modeling. The ability to infer parameters via an inverse problem allows practitioners to gain important information about systems, processes and the models that are used to describe them. Such inference also avoids intrusive experiments, which may otherwise be prohibitively expensive or excessively intrusive. Once parameters have been inferred via an inverse problem, they can be used for forward propogation of the model and computing relevant goals and quantities of interest for the modeling/decision making process.

Data, constraints and prior information form the critical components of an inverse problem formulation. There is an extensive literature pertaining to all these components, pertaining to the ascription/filtering of noise in data, the choice of regularization/prior information, the efficient incorporation of constraints in solving the problem and so on. Specifically, the work on the constraints or `state equations' in the inverse problem focuses on methods to reduce the computational expense of solving/inverting the models that represent them. Such work includes the use of reduced order models to represent constraints, adaptive mesh refinement to control the size of the state equations among other strategies.

Models that represent phenomena and processes can usually be organized in a hierarchy according to their fidelity and expense. The use of multiple models simultaneously for forward simulation is well established. In this work, we extend the use of the multi-fidelity models to inverse problems. We present a systematic method to use different constraints in different regions of the domain, while controlling the error due to the use of lower fidelity models in a target quantity of interest. Our approach is based on a novel use of adjoint error estimation, which enables the computation of third-order QoI error estimates, at the cost of solving an additional adjoint problem.

>>>>

Parameter estimation/inference is an important component of mathematical modeling. The ability to infer parameters via an inverse problem allows practitioners to gain important information about systems, processes and the models that are used to describe them. Such inference also avoids intrusive experiments, which may otherwise be prohibitively expensive or impractical. Once parameters have been inferred via an inverse problem, they can be used for forward propagation of the model and, computing relevant quantities of interest for the modeling/decision making process.

Data, constraints and prior information form the critical components of an inverse problem formulation. There is an extensive literature pertaining to all these components, including the filtering of noise in data, the choice of regularization/prior information, the efficient incorporation of constraints in solving the problem and so on. Specifically, the work on the constraints or `state equations' in the inverse problem focuses on reducing the computational expense of solving/inverting the models that represent them. Such work includes the use of reduced order models to represent constraints~\cite{}, adaptive mesh refinement to control the size of the state equation discretizations~\cite{} among other strategies.

Models that represent phenomena and processes can usually be organized in a hierarchy according to their fidelity and expense. The use of multiple models simultaneously for forward simulation is well established. In this work, we extend the use of multi-fidelity models to inverse problems. We present a systematic method to use different constraints in different regions of the domain, while controlling the error due to the use of lower fidelity models, in a target quantity of interest. Our approach is based on a novel use of adjoint error estimation, which enables the computation of third-order QoI error estimates, at the cost of solving an additional adjoint problem.

### 2016-05-22 paper/mathematical_formulation.tex
Minor typo
Minor typos (more to come)

%------------------------------------------------------------------------------%
\subsection{Problem Setup}  \label{sec:setup}
%------------------------------------------------------------------------------%
%
We seek to infer an unknown parameter $q \in Q$, where $Q$ is a Hilbert space. Consider a state variable $u \in U$, which satisfies a bilinear form $a:U \times U \to \Reals$, such that,
%
\begin{equation}
\label{eq:weakForm}
a(u,q)(\phi)=\ell(q)(\phi),\quad\forall\phi\in U,
\end{equation}
%
where the form $a$ and functional $\ell$ are linear with respect to the arguments in the second pair of parentheses. Further, we define an observation operator $C:U\to\Reals^{n_d}$ that maps the state to $n_d$ predicted observations. The actual observations (data) are denoted by $d\in\R^{n_d}$.

The unknown parameter $q$ can then be inferred by minimizing the difference between the predicted and actual observations, leading to an inverse problem. Such inverse problems are typically ill-posed, since the observations are sparse, and insufficiently informative to uniquely determine the parameters. To make the inverse problem well-posed, a regularization, denoted by $R(q)$, is used to inject prior information or beliefs about the parameters into the formulation. The regularized inverse problem can thus be written as a constrained optimization problem,
%
\begin{subequations}
\label{eq:invOpt}
\begin{align}
\min\limits_{q,u} & \quad J(q,u)=\frac{1}{2}\|d-C(u)\|_2^2 + R(q) \label{eq:invOpt_obj} \\
\textrm{s.t. }& \quad a(u,q)(\phi)=\ell(q)(\phi),\quad\forall\phi\in U \label{eq:invOpt_cons}
\end{align}
\end{subequations}
%
where we aim to minimize the cost function $J$, which includes the mismatch between predicted and actual observations and a regularization term $R(q)$, subject to the state $u$ and parameters $q$ satisfying the constraints given by Equation (\ref{eq:weakForm}).

The constraints given by Eq.~\eqref{eq:weakForm} are typically models of physical processes or systems. However, a given physical system need not have a unique model that can describe it; there may be various models of different fidelities. This naturally introduces a hierarchy of models, introducing a tradeoff between fidelity and computational expense. Consider two models with which we infer parameters: the high-fidelity (HF) model and a low fidelity (LF) model. We then have a specific form of Equation (\ref{eq:weakForm}) for the high-fidelity model:
\begin{equation}
a_{HF}(u_{HF},q_{HF})(\phi_{HF})=\ell_{HF}(q_{HF})(\phi_{HF}),\quad\forall\phi_{HF}\in U_{HF},
\end{equation}
where $u_{HF}\in U_{HF}$ and $q_{HF}\in Q_{HF}$. Similarly, the inverse problem in Equation (\ref{eq:invOpt}) has the specific form
\begin{equation}
\begin{array}{r@{}c}
\min\limits_{q_{HF},u_{HF}} & \quad J_{HF}(q_{HF},u_{HF})=\frac{1}{2}\|d-C_{HF}(u_{HF})\|_2^2 + R_{HF}(q_{HF}) \\ \textrm{s.t. }& \quad a_{HF}(u_{HF},q_{HF})(\phi_{HF})=\ell_{HF}(q_{HF})(\phi_{HF}),\quad\forall\phi_{HF}\in U_{HF},
\end{array}
\end{equation}
for which we have the Lagrangian
\begin{equation}
\mathcal{L}_{HF}(q_{HF},u_{HF},z_{HF})= J_{HF}(q_{HF},u_{HF})-(a_{HF}(u_{HF},q_{HF})(z_{HF})-\ell_{HF}(q_{HF})(z_{HF})).
\end{equation}
The lower fidelity model can be similarly defined and identified by the subscript $LF$.

@@

Extending the property in Equation (\ref{eq:MeqI}) to the augmented Lagrangians for the high and low fidelity models, we have,
%
\begin{multline}
\label{eq:repIwithM}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\\mathcal{M}_{HF}(\Psi_{HF})-\mathcal{M}_{HF}(\Psi_{LF})+\mathcal{M}_{HF}(\Psi_{LF})-\mathcal{M}_{LF}(\Psi_{LF})\textrm{.}
\end{multline}
%
Applying Proposition 3 from~\cite{BecVex05} for the difference $\mathcal{M}_{HF}(\Psi_{HF})-\mathcal{M}_{HF}(\Psi_{LF})$,  
\begin{equation}
\mathcal{M}_{HF}(\Psi_{HF})-\mathcal{M}_{HF}(\Psi_{LF}) = \frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{LF})(\Psi_{HF}-\Psi_{LF})+\mathcal{R}(e^3)\textrm{,}
\end{equation}
where $\mathcal{R}$ is a remainder term that is third-order in the error $e=\Psi_{HF}-\Psi_{LF}$. Combining Equations (\ref{eq:repIwithM}) and (\ref{eq:preadj}) we obtain
\begin{multline}
\label{eq:preadj}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\\frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{LF})(\Psi_{HF}-\Psi_{LF})+\mathcal{M}_{HF}(\Psi_{LF})-\mathcal{M}_{LF}(\Psi_{LF})+\mathcal{R}(e^3)\textrm{.}
\end{multline}

@@

Combining Equations (\ref{eq:preadj}) and (\ref{eq:adjOutErr}), we have,
\begin{multline}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\-\frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{LF})(\Lambda)+\mathcal M_{HF}(\Psi_{LF})-\mathcal M_{LF}(\Psi_{LF})+\mathcal{R}(e^3).
\end{multline}
%
which completes the proof.
\end{proof}
%
The above error estimate is general and the lower-fidelity model can also be a mixed-fidelity model that combines the high- and low-fidelity model. Given a low-fidelity model and a high-fidelity model, an intermediate, mixed-fidelity (MF) model can be formed by using the high-fidelity model in some parts of the domain, and the low-fidelity model in the rest of the domain.

>>>>

%------------------------------------------------------------------------------%
\subsection{Problem Setup}  \label{sec:setup}
%------------------------------------------------------------------------------%
%
Let $Q$, $U$ be Hilbert spaces. We seek to infer an unknown parameter $q \in Q$. The parameter is constrained by a state equation, with a state variable $u \in U$, which satisfies a bilinear form $a:U \times U \to \Reals$, such that,
%
\begin{equation}
\label{eq:weakForm}
a(u,q)(\phi)=\ell(q)(\phi),\quad\forall\phi\in U.
\end{equation}
%
The bilinear form $a$ and functional $\ell$ are linear with respect to the arguments in the second pair of parentheses. Further, we define an observation operator $C:U\to\Reals^{n_d}$ that maps the state to $n_d$ predicted observations. The actual observations (data) are denoted by $d\in\R^{n_d}$.

The unknown parameter $q$ can then be inferred by minimizing the difference between the predicted and actual observations, leading to an inverse problem. Such inverse problems are typically ill-posed, since the observations are sparse, and insufficiently informative to uniquely determine the parameters. To make the inverse problem well-posed, a regularization, denoted by $R(q)$, is used to inject prior information or beliefs about the parameters into the formulation. The regularized inverse problem can thus be written as a constrained optimization problem,
%
\begin{subequations}
\label{eq:invOpt}
\begin{align}
\min\limits_{q,u} & \quad J(q,u)=\frac{1}{2}\|d-C(u)\|_2^2 + R(q), \label{eq:invOpt_obj} \\
\textrm{s.t. }& \quad a(u,q)(\phi)=\ell(q)(\phi),\quad\forall\phi\in U. \label{eq:invOpt_cons}
\end{align}
\end{subequations}
%
Thus, we aim to minimize the cost function $J$, which includes the mismatch between predicted and actual observations and a regularization term $R(q)$, subject to the state $u$ and parameters $q$ satisfying the constraints given by Equation (\ref{eq:weakForm}).

The constraints given by Eq.~\eqref{eq:weakForm} are typically models of physical processes or systems. However, a given physical system need not have a unique model that can describe it; there may be various models of different fidelities. This naturally introduces a hierarchy of models and, a tradeoff between fidelity and computational expense. Consider two models with which we infer parameters: a high-fidelity (HF) model and a low fidelity (LF) model. We then have a specific form of the inverse problem in Equation (\ref{eq:invOpt}) with this model,
\begin{equation}
\begin{array}{r@{}c}
\min\limits_{q_{HF},u_{HF}} & \quad J_{HF}(q_{HF},u_{HF})=\frac{1}{2}\|d-C_{HF}(u_{HF})\|_2^2 + R_{HF}(q_{HF}) \\ \textrm{s.t. }& \quad a_{HF}(u_{HF},q_{HF})(\phi_{HF})=\ell_{HF}(q_{HF})(\phi_{HF}),\quad\forall\phi_{HF}\in U_{HF},
\end{array}
\end{equation}
for which we have the Lagrangian,
\begin{multline}
\mathcal{L}_{HF}(q_{HF},u_{HF},z_{HF}) = \\
J_{HF}(q_{HF},u_{HF})-(a_{HF}(u_{HF},q_{HF})(z_{HF})-\ell_{HF}(q_{HF})(z_{HF})).
\end{multline}
The inverse problem formulation and Lagrangian with the lower fidelity model can be similarly defined, and be identified with subscript $LF$.

@@

Extending the property in Equation (\ref{eq:MeqI}) to the augmented Lagrangians for the high and low fidelity models, we have,
%
\begin{multline}
\label{eq:repIwithM}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\\mathcal{M}_{HF}(\Psi_{HF})-\mathcal{M}_{HF}(\Psi_{LF})+\mathcal{M}_{HF}(\Psi_{LF})-\mathcal{M}_{LF}(\Psi_{LF})\textrm{.}
\end{multline}
%
Applying Proposition 3 from~\cite{BecVex05} for the difference $\mathcal{M}_{HF}(\Psi_{HF})-\mathcal{M}_{HF}(\Psi_{LF})$,
\begin{equation}
\label{eq:beckvex}
\mathcal{M}_{HF}(\Psi_{HF})-\mathcal{M}_{HF}(\Psi_{LF}) = \frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{LF})(\Psi_{HF}-\Psi_{LF})+\mathcal{R}(e^3)\textrm{,}
\end{equation}
where $\mathcal{R}$ is a remainder term that is third-order in the error $e=\Psi_{HF}-\Psi_{LF}$. Combining Equations (\ref{eq:repIwithM}) and (\ref{eq:beckvex}) we obtain
\begin{multline}
\label{eq:preadj}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\\frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{LF})(\Psi_{HF}-\Psi_{LF})+\mathcal{M}_{HF}(\Psi_{LF})-\mathcal{M}_{LF}(\Psi_{LF})+\mathcal{R}(e^3)\textrm{.}
\end{multline}

@@

Combining Equations (\ref{eq:preadj}) and (\ref{eq:adjOutErr}), we have,
\begin{multline}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\-\frac{1}{2}\mathcal{M}'_{HF,\Psi}(\Psi_{LF})(\Lambda)+\mathcal M_{HF}(\Psi_{LF})-\mathcal M_{LF}(\Psi_{LF})+\mathcal{R}(e^3). \nonumber
\end{multline}
%
which completes the proof.
\end{proof}
%
The above error estimate is general and the lower-fidelity model can also be a mixed-fidelity model that combines the high- and low-fidelity model. Given a low-fidelity model and a high-fidelity model, an intermediate, mixed-fidelity (MF) model can be formed by using the high-fidelity model in some parts of the domain, and the low-fidelity model in the rest of the domain.
